---
title: "Chapter 1 Supplement"
output:
  html_document:
    df_print: paged
---

```{r setup}
library(tidyverse)
library(zoo)
library(deSolve)
library(ggforce)
```

# Measles Notifications in New York City, 1950-1960

```{r,include=FALSE}
# knitr::opts_chunk$set(echo = FALSE,fig.height=2.5,fig.width=7)

new_york_measles <- read.csv("new_york_measlescsv.csv")

new_york_population <- read.csv("population_stats_NYC.csv")

# new_york_year_avg <- read.csv("new_york_year_avg.csv")
```

Filtering data to only include the range from 1950 to 1960:

```{r}
new_york_measles <- new_york_measles %>%
  filter(numdate >= 1950, numdate <=1960)

new_york_population <- new_york_population %>%
  filter(year >= 1950, year<=1960)
```

Graphs for population data:

```{r}
new_york_population %>%
  mutate(percent.pop.change = dpop/pop,birth.rate = births/pop,death.rate = deaths/pop) %>%
  select(-infmort,-vaccprop,-pop,-births,-deaths) %>%
  pivot_longer(cols = c("percent.pop.change","birth.rate","death.rate"),names_to = c("Series"),values_to = "Number") %>%
  mutate(Series = fct_relevel(Series,c("percent.pop.change","birth.rate","death.rate"))) %>%
  ggplot(aes(x=year,y=Number)) +
  geom_line(aes(color = Series)) +
  facet_grid(Series ~ .,scales = "free") +
  labs(title = "New York Population Data 1950-1960",x="Year",y="Rate") +
  scale_color_manual(values = c("#77c298","#718dbf","#e84d60"),labels = c("Fraction population change","Birth rate per capita","Death rate per capita")) +
  theme_bw() +
  theme(strip.text.y = element_blank(),legend.position = "bottom")
```

Graph for of weekly notifications:

```{r}
new_york_year_avg <- new_york_measles %>%
  mutate(year = floor(numdate)) %>%
  group_by(year) %>%
  summarise(cases = mean(cases)) %>%
  mutate(Series = "Summary")

colnames(new_york_year_avg)[1] <- "Date"

ny_measles_plot <- new_york_measles %>%
  select(numdate,cases) %>%
  mutate(Series = "Weekly Notifications")
  
colnames(ny_measles_plot)[1] <- "Date"

ny_measles_plot <- rbind(new_york_year_avg,ny_measles_plot)

ggplot() +
  geom_line(data = new_york_measles,aes(x=numdate,y=cases,color="Weekly Notifications")) +
  geom_step(data = new_york_year_avg,aes(x=Date,y=cases,color="Average Weekly notifications")) +
  labs(title =" Notifications of Measles Cases in New York City, 1950-1960",x= "Year",y="Number of Notifications") +
scale_color_manual(values = c("#718dbf","#e84d60")) +
  theme_bw() +
  theme(legend.position = "bottom")
```

Proportion of NYC population infected boxplot:

```{r}
ny_measles_prevalence <- new_york_measles %>%
  mutate(year = floor(numdate))

ny_measles_prevalence <- merge(ny_measles_prevalence,new_york_population,by="year")

ny_measles_prevalence <- ny_measles_prevalence %>%
  mutate(prop.infected = cases/pop,pattern = ifelse(year %% 2 == 0,"High year","Low year"))

ny_measles_prevalence %>%
  mutate(month = as.factor(floor((numdate-year)*12)+1)) %>% 
  ggplot(aes(month,prop.infected)) +
  geom_boxplot(aes(group = month)) +
  labs(title="Proportion of NYC Population Infected by Month, 1950-1960",x="Month",y="Proportion of Population Infected") +
  facet_wrap(pattern ~ .) +
  theme_bw()
```

Average total cases in high years and low years:

````{r}
high_low <- new_york_measles %>%
  mutate(numdate = floor(numdate)) %>%
  group_by(numdate) %>%
  summarise(total_cases = sum(cases)) %>%
  mutate(high_low = ifelse(numdate %% 2 == 0,"High","Low")) %>%
  group_by(high_low) %>%
  summarise(total_cases = mean(total_cases))

high_low
```

Ratio of total cases in high years vs low years:

```{r}
high_low$total_cases[1]/high_low$total_cases[2]
```

Size of peaks in high and low years:

```{r}
high_low2 <- new_york_measles %>%
  mutate(numdate = floor(numdate)) %>%
  group_by(numdate) %>%
  summarise(max_cases = max(cases)) %>%
  mutate(high_low = ifelse(numdate %% 2 == 0,"High","Low")) %>%
  group_by(high_low) %>%
  summarise(max_cases = mean(max_cases))

high_low2
```

Ratio of peak cases in high years vs low years:

```{r}
high_low2$max_cases[1]/high_low2$max_cases[2]
```

